Measuring background and systemic risks in finance

### 2015-06-04 srisk5.tex
pdj changes

% Title of paper
\title{Measuring background and systemic risks in finance}
% List of authors, with corresponding author marked by asterisk
\author{Piet de Jong,  Geoff Loudon and Weihao Choo \\[4pt]
% Author addresses
\textit{Department of Applied Finance and Actuarial Studies\\ Macquarie University, Sydney, NSW 2109.}
\\[2pt]
%E-mail address for correspondence
{piet.dejong@mq.edu.au}}

@@

\begin{abstract}
This article refines, builds on, and extends  SRISK methodology recently proposed in the literature.  The refinement is to define SRISK in terms of a put on the Basel shortfall.   This  is built on by defining the background and systemic stress  of a firm as unconditional and departure from unconditional  expectation of the put, the latter when a hypothetical systemic stress is applied.  Systemic stress is defined in terms of a random variable and can take on variety of forms including alternative scenarios in usual stress testing as well stress driven by the interaction of variables.  Stressor random variables  are chosen by the practitioner.  Stressed expectations are linear, a sector systemic stress is naturally defined as linear in the firm specific systemic stress.      Application is made to Australian financial data.
\end{abstract}

@@

JEL classification
C15; E53; G21  (Needs to be filled out)

\section{Background}

Recent literature in the area of quantitative systemic risk or stress measurement includes
\cite{adrian2011covar},
\cite{acharya2012capital},
\cite{acharya2012measuring},
 \cite{brownlees2010volatility} and \cite{brownlees2015}.   This study aims to extend and apply these techniques particularly in relation to the entities regulated by the Australian Prudential Regulation Authority (APRA). Thus our  broad aim is to develop, implement and bring to bear recent developments in systemic risk measurement and stress testing  on the aims of APRA and the CIFR targeted research areas.   (This needs much more commentary).

This study refines and develops recently proposed methods and applies the same to publicly available daily financial data for  the eight  Australian banks detailed in \tref{eightbanks} spanning the period from   3 April 2000 through to 1 December 2014.  The data is described in detail in \aref{data}.  Prices, adjusted for dividends are plotted in \fref{prices}.   Prices are combined with number of shares to derive total equity.    Total debt for each firm at each day is also collected.  (Should these be plotted?).

Remaining sections are structured as follows. \sref{capshort} and \sref{futshort} define current and future  capital shortfall similarly as \cite{brownlees2015} under a simplistic Basel II regulatory framework. The definition of systemic risk used by \cite{brownlees2015} is discussed in \sref{srisk}, and is refined, extended and illustrated \sref{improve} to \sref{estimate}. \sref{marketstress} discusses specific stress examples which extend beyond \cite{brownlees2015}. Applications to Australian bank data are discussed in \sref{simulate} and \sref{simulate1}. \sref{aggregate} and \sref{aggregate1} discuss how stresses are combined across firms with and without allowance for merging or diversification. Generalised stressed are outlined in \sref{genstress}. \sref{riskmethod} offers an alternative modeling approach, based on assets. \sref{comparison} summarises differences between proposed stress measurement in this paper with that of 
\cite{brownlees2015}. \sref{conclude} concludes.

@@

@@

 \cite{brownlees2015} defines  systemic risk (SRISK) for a group of firms  at time $t$ as
\be{esrisk}
\mathrm{SRISK}_t\equiv\kappa\sum_i d_{it}\left|1-\Es(\e^{\nu_{it}-\ell_{it}})\right|^+ \ ,
\ee
where  $\Es$ denotes expectation given a major general market downturn. Hence the expected capital shortfall (allowing surplus offset) is computed from \eref{shortfall} assuming a market downturn and added across all firms. Firms having a expected capital surplus are ignored. The systemic risk of firm $i$ at time $t$ is its proportionate contribution to \eref{esrisk}:
 \be{sriskperc}
 \mathrm{SRISK}_{it}\equiv \pi_{it}\frac{|1-\Es(\e^{\nu_{it}-\ell_{it}})|^+}{\Ex_d\{|1-\Es(\e^{\nu_{it}-\ell_{it}})|^+\}}\cq \pi_{it}\equiv \frac{d_{it}}{d_t}\cq d_t\equiv \sum_i d_{it}\ ,
 \ee
where $\Ex_d$ denotes weighted averaging across firms $i$ using weights $\pi_{it}$.     Large $\mathrm{SRISK}_{it}$ indicates firm $i$ is systemically important: it holds a high proportion of the total debt, and it is likely to heavily breach the Basel capital requirement compared to remaining other firms.  Expression \eref{sriskperc} depends on $\kappa$ through each of the $\ell_{it}$.

@@

\section{Systemic beta's}

@@

The expressions in \eref{zscore2} suggests two ways of  thinking about the  systemic beta's  $\beta_{it}$.   First, the left hand side equation shows systemic stress $\beta_{it}$ serves to shift the mean $\E(p_{it})$ of put prices  by, on average, $\sigma_\phi$.    The stress from $\phi$ is captured with $\sigma_\phi\beta_{it}$, the second term on the right of the first expression.   This term is called systemic stress and represents the addition to expected put prices if stress $\phi$ materialises.

Systemic beta's are stated in terms of additions to the normal put price $\E(p_{it})$. Thus in an increasingly dire financial situation $\E(p_{it})$ will increase.   However  $\beta_{it}$ and the systemic stress measures the  change in current put values  under  further $\phi$--stress.  This varies from  \cite{brownlees2015} where $\mathrm{SRISK}_i$ does not distinguish between the current, possibly high, put prices $\E(p_{it})$ and the potential  increment due to potential stress.   Thus in an increasingly dire financial situation put prices are liable to increase on account of decreasing expected returns and increasing volatility.   However our definition of systemic stress using systemic beta's measures the further effect on account of potential extra stress imposed onto the system over and above the already existing stress.   Systemic beta's capture   ``marginal" effects.

@@

The estimation of background stresses and systemic stresses and beta's requires projections of future capital shortfalls. As in \cite{brownlees2015}, projections in this paper are constructed  using time series models of forward rates of return $\nu_{it}$ for firm $i$ and $\nu_{mt}$ for the market. The market return is used as the stressor with different choices of the stress function $\phi$ modelling different stress scenarios.

@@

(Geoff -- can you fill out?)

>>>>

% Title of paper
\title{Measuring background and systemic risks in finance}
% List of authors, with corresponding author marked by asterisk
\author{Piet de Jong,  Geoffrey loudon and Weihao Choo \\[4pt]
% Author addresses
\textit{Department of Applied Finance and Actuarial Studies\\ Macquarie University, Sydney, NSW 2109.}
\\[2pt]
%E-mail address for correspondence
{piet.dejong@mq.edu.au}}

@@

\begin{abstract}
This article refines, builds on, and extends the systemic risk methodology [SRISK] recently proposed by \cite{brownlees2015}. We refine SRISK by defining it in terms of a put on the Basel shortfall. We build on this by distinguishing background and systemic stress of a firm. Background stress is the unconditional expectation of the put. Systemic stress is the departure from the unconditional expectation when a hypothetical systemic stress is applied.  Systemic stress is defined in terms of a random variable and can take on variety of forms including alternative scenarios as in usual stress testing and also 
stress driven by the interaction of variables. Stressor random variables are chosen by the practitioner. Stressed expectations are linear and sector systemic stress is naturally defined as linear in the firm specific systemic stress. Application is made to Australian financial data.
\end{abstract}

@@

JEL classification
C15; E53; G21  %(Needs to be filled out)

\section{Introduction}

Monitoring systemic risk is a key function of macro-prudential regulation. Central banks and other financial regulatory bodies such as the US Financial Stability Oversight Council, the European Systemic Risk Board and the Australian Prudential Regulation Authority [APRA] are concerned with developing and applying quantitative risk measures to alert them of potential episodes of heightened systemic risk. Systemic risk has many dimensions. For example, in their influential review paper, \cite{Bisias2012} present a taxonomy of six categories of systemic risk comprising 31 quantitative risk measures. While attempts to gain consensus on a theoretical definition have proved elusive, several empirical measures of systemic risk have been developed and applied with some success.  An important example is the SRISK measure due to \cite{brownlees2015} which proposes an index that measures the expected capital shortfall on a financial institution conditional on a large and lengthy decline in equity markets. \cite{brownlees2015} show the practical usefulness of their measure in three ways: (i) SRISK rankings identify systemically risky US banks during the global financial crisis [GFC]; (ii) pre-crisis SRISK helps predict capital injections by the Fed Reserve; (iii) aggregate SRISK provides early warning of declines in industrial production and higher unemployment.

Our paper builds upon and extends the model of \cite{brownlees2015} in several interesting and important ways. First, we reinterpret their SRISK measure as a put on the expected shortfall using stressed expectation. While they condition on the event of a severe market downturn, we take the stressed expectation of the put value. This enables us to capture the impact of increased return volatility due to stress. Second, we decompose their SRISK measure into background stress and imposed system stress. Background stress is a function of existing volatility and leverage. As these factors increase, background stress rises and this makes the firm more susceptible to systemic risk, regardless of the stress event itself. Third,    
\cite{brownlees2015} condition on an absolute cutoff of arbitrary size and length. We suggest that it is also helpful to specify conditional behaviour in percentile terms relative to current market conditions. This provides a sharper focus on the incremental impact of further stress and allows us to draw stronger implications of SRISK numbers. These extensions provide additional insights into the sources and economic nature of systemic risk. They provide potentially new and improved ways of anticipating episodes of heightened systemic risk. Their practical use can be expected to help regulators monitor systemic risk in a timely fashion to facilitate remedial action where necessary.

% This study aims to extend and apply these techniques particularly in relation to the entities regulated by the Australian Prudential Regulation Authority (APRA). Thus our  broad aim is to develop, implement and bring to bear recent developments in systemic risk measurement and stress testing  on the aims of APRA and the CIFR targeted research areas.   (This needs much more commentary).

While our measures are designed for use in all financial systems, it is of particular interest to test them using the Australian setting for at least two reasons. First, the Australian financial system is dominated by only four banks. In Australia there are four major banks that hold 78\% of the total assets of all Authorised Deposit-taking Institutions in  Australia.\footnote{Based on balances at 31 December 2014 disclosed in APRA's Quarterly Authorised Deposit-taking Institution [ADI] Performance Report for March 2015.} It is therefore obvious that the four majors are systemically important and the remaining ADIs are not so unless stress in one of these minors somehow creates financial contagion to the majors, in an unforeseeable way. Second, the Australian banking system proved to be resilient through the recent GFC, implying that systemic risk did not reach the extreme levels experienced by other countries where bank failures occurred. Both these features of the Australian system suggest that it may be difficult to find useful additional information about systemic risk by refining quantitative measures like SRISK. Since we are able to show that our measures are informative about systemic risk in Australia, we conjecture that they will provide additional insights elsewhere.

To illustrate the nature and usefulness of our measures, we use publicly available daily financial data for the eight Australian banks detailed in \tref{eightbanks} spanning the period from  3 April 2000 through to 1 December 2014.  The data is described in detail in \aref{data}.  Prices, adjusted for dividends are plotted in \fref{prices}.  Prices are combined with number of shares to derive total equity. Total debt for each firm is also collected from annual accounting reports. % (Should these be plotted? No, not informative;  leverage ratio plots are sufficient).

Remaining sections are structured as follows. \sref{litrev} briefly discusses some related literature. \sref{capshort} and \sref{futshort} define current and future  capital shortfall similarly as \cite{brownlees2015} under a simplistic Basel II regulatory framework. The definition of systemic risk used by \cite{brownlees2015} is discussed in \sref{srisk}, and is refined, extended and illustrated \sref{improve} to \sref{estimate}. \sref{marketstress} discusses specific stress examples which extend beyond \cite{brownlees2015}. Applications to Australian bank data are discussed in \sref{simulate} and \sref{simulate1}. \sref{aggregate} and \sref{aggregate1} discuss how stresses are combined across firms with and without allowance for merging or diversification. Generalised stressed are outlined in \sref{genstress}. \sref{riskmethod} offers an alternative modeling approach, based on assets. \sref{comparison} summarises differences between proposed stress measurement in this paper with that of 
\cite{brownlees2015}. \sref{conclude} concludes.

@@

\section{Literature review}\label{litrev}
Many papers in the finance and economics literature are devoted to the development and application of a variety of quantitative measures of systemic risk or financial stress. \cite{Bisias2012} provide an instructive and qualitative survey of numerous measures in use. \cite{Giglio2015} supply empirical evidence on the ability of many of these measures, both individually and collectively, to provide early warning signals of deterioration in macro-economic conditions. In this section, we provide a selective overview of the measures most closely related to SRISK and its relevance for assessing systemic risk in the financial sector.

The genesis of SRISK lies in a series of papers including \cite{acharya2012capital}, \cite{acharya2012measuring} and \cite{brownlees2010volatility}. In these papers, SRISK is defined as the expected capital shortfall of a financial firm, conditional on a crisis. A crisis is deemed to occur whenever the relevant market index suffers a significant decline over a chosen horizon. \cite{Benoit2013} show how several popular systemic risk measures including MES and SRISK are transformations of market risk measures and derive conditions under which these measures provide similar rankings of systemically important financial institutions.

\cite{brownlees2015} develop SRISK. [***** to add a fuller description of this paper given it is the key paper for our work]

\cite{Engle2015} extend the model in \cite{brownlees2015} to incorporate global, country and industry effects. This extended model makes effective use of \cite{Engle2014dcb} dynamic conditional model. Their empirical methods are able to provide information of the relative systemic importance of industry type and country identity.

\cite{adrian2011covar} propose an alternative measure of systemic risk that emphasises the co-dependence of financial firms and the importance of risk spillovers within the financial sector. They introduce CoVaR$_i$ which computes the Value-at-Risk [VaR] of the entire financial system, conditional on institution $i$ being in distress. Distress is defined as the state where the institution is exactly at its VaR level. \cite{Girardi2013} improve CoVaR by conditioning on the institution being at most at its VaR. This generalisation is useful as it takes into account severity of tail losses and facilitates back-testing of CoVaR.

Empirical estimates of SRISK measure require a model for the joint distribution of returns of individual firms and the market. Many of these papers use \cite{engle2002dynamic}'s GARCH-DCC model which nicely captures the time-varying dynamics of the return (co)variances in a feasible manner. While we also use this model in our empirical analysis, we emphasize that our approach can utilise any appropriate model for simulating forward the joint distribution of firm and market returns.

@@

 \cite{brownlees2015} defines  systemic risk (SRISK) for a group of firms  at time $t$ as
\be{esrisk}
\mathrm{SRISK}_t\equiv\kappa\sum_i d_{it}\left|1-\Es(\e^{\nu_{it}-\ell_{it}})\right|^+ \ ,
\ee
where  $\Es$ denotes expectation given a major general market downturn. Hence the expected capital shortfall (allowing surplus offset) is computed from \eref{shortfall} assuming a market downturn and added across all firms. Firms having an expected capital surplus are ignored. The systemic risk of firm $i$ at time $t$ is its proportionate contribution to \eref{esrisk}:
 \be{sriskperc}
 \mathrm{SRISK}_{it}\equiv \pi_{it}\frac{|1-\Es(\e^{\nu_{it}-\ell_{it}})|^+}{\Ex_d\{|1-\Es(\e^{\nu_{it}-\ell_{it}})|^+\}}\cq \pi_{it}\equiv \frac{d_{it}}{d_t}\cq d_t\equiv \sum_i d_{it}\ ,
 \ee
where $\Ex_d$ denotes weighted averaging across firms $i$ using weights $\pi_{it}$.     Large $\mathrm{SRISK}_{it}$ indicates firm $i$ is systemically important: it holds a high proportion of the total debt, and it is likely to heavily breach the Basel capital requirement compared to remaining other firms.  Expression \eref{sriskperc} depends on $\kappa$ through each of the $\ell_{it}$.

@@

\section{Systemic betas}

@@

The expressions in \eref{zscore2} suggests two ways of  thinking about the  systemic betas  $\beta_{it}$.   First, the left hand side equation shows systemic stress $\beta_{it}$ serves to shift the mean $\E(p_{it})$ of put prices  by, on average, $\sigma_\phi$.    The stress from $\phi$ is captured with $\sigma_\phi\beta_{it}$, the second term on the right of the first expression.   This term is called systemic stress and represents the addition to expected put prices if stress $\phi$ materialises.

Systemic betas are stated in terms of additions to the normal put price $\E(p_{it})$. Thus in an increasingly dire financial situation $\E(p_{it})$ will increase.   However  $\beta_{it}$ and the systemic stress measures the  change in current put values  under  further $\phi$--stress.  This varies from  \cite{brownlees2015} where $\mathrm{SRISK}_i$ does not distinguish between the current, possibly high, put prices $\E(p_{it})$ and the potential  increment due to potential stress.   Thus in an increasingly dire financial situation put prices are liable to increase on account of decreasing expected returns and increasing volatility.   However our definition of systemic stress using systemic betas measures the further effect on account of potential extra stress imposed onto the system over and above the already existing stress.   Systemic betas capture   ``marginal" effects.

@@

The estimation of background stresses and systemic stresses and betas requires projections of future capital shortfalls. As in \cite{brownlees2015}, projections in this paper are constructed  using time series models of forward rates of return $\nu_{it}$ for firm $i$ and $\nu_{mt}$ for the market. The market return is used as the stressor with different choices of the stress function $\phi$ modelling different stress scenarios.

@@

Data are sourced from DataStream. DataStream company codes are 
\begin{center}
	\begin{tabular}{ll}
A:ANZX & Australia and New Zealand Banking Group Limited\\
A:ABAX & Auswide Bank Limited\\
A:BOQX & Bank of Queensland Limited\\
A:BENX & Bendigo and Adelaide Bank Limited\\
A:CBAX & Commonwealth Bank of Australia\\
A:MQGX & Macquarie Bank Limited\\
A:NABX & National Australia Bank Limited\\
A:WBCX & Westpac Banking Corporation.\\	
\end{tabular}
\end{center}

DataStream datatype codes are
\begin{center}
	\begin{tabular}{ll}
	NOSH & Number of Shares\\
	P & Price\\
	WC02999 & Total Assets\\
	WC03501 & Shareholders Equity\\
	RZ & Total return index\\
\end{tabular}
\end{center}

Firm equity $w_{it}$ is Number of Shares multiplied by Price. Firm debt $d_{it}$ is Total Assets minus Shareholders Equity. Stock returns are computed from the Total Return Index which includes dividends as paid.
